

Just loading any necessary libraries

```{r libraries}
library(tidyverse)
library(ggplot2)
```


Our Question: Would a population with a higher r value but lower K or a population with a lower
r value but a higher K would be able to expand there range faster across habitat patches.

We hypothesized that the population with higher r but lower K would be able to expand there range 
at a faster rate.

In order to build the simulation we had to make make several assumptions about the population:

-Migration is a random event

-Individuals can only migrate to the nearest patch on either side

-Constant growth rate r

-Carrying capacity (K) is fixed for each population

-No difference in the quality of habitat patches

-The populations are only moving in 1 dimensional space

-No competition between species

'Wins' were counted based on what population reached the furthest patch from the starting patch
first.


Nested list containing all the patches and the populations for both n1 and n2 inside. With a separate 
nested list containg the parameters for both populations.
```{r worlds and params generator}
my_world <- list("1" = list("n1" = 10, "n2" = 10), 
              "2" = list("n1" = 0, "n2" = 0), 
              "3" = list("n1" = 0, "n2" = 0),
              "4" = list("n1" = 0, "n2" = 0), 
              "5" = list("n1" = 0, "n2" = 0), 
              "6" = list("n1" = 0, "n2" = 0), 
              "7" = list("n1" = 0, "n2" = 0), 
              "8" = list("n1" = 0, "n2" = 0), 
              "9" = list("n1" = 0, "n2" = 0), 
              "10" = list("n1" = 0, "n2" = 0),
              "11" = list("n1" = 0, "n2" = 0),
              "12" = list("n1" = 0, "n2" = 0),
              "13" = list("n1" = 0, "n2" = 0),
              "14" = list("n1" = 0, "n2" = 0),
              "15" = list("n1" = 0, "n2" = 0),
              "16" = list("n1" = 0, "n2" = 0),
              "17" = list("n1" = 0, "n2" = 0),
              "18" = list("n1" = 0, "n2" = 0),
              "19" = list("n1" = 0, "n2" = 0),
              "20" = list("n1" = 0, "n2" = 0))
my_params <- list("n1" = list("r" = 0.3, "k" = 300), 
               "n2" = list("r" = 0.6, "k" = 150))
```
Now that the world is created, we  made a set of functions and algorithms to run our simulation

The migration function uses a random binomial draw to determine the number of migrants in a 
population
```{r migrants generator}
migrants_number <- function(n){
  return(rbinom(n = 1, size = n, prob = 0.5))
}
```

The direction generator function random picks between 1 and -1 to determine what direction the
migrants will move.
```{r direction generator function}
direction_generator <-function(){
  return(sample(c(-1, 1), 1))
}
direction_generator()

```

Our popultion growth function uses a discrete time logsistic growth equation, that produces the 
population in the next time step instead of returning a rate of change.
```{r growth function}
growth <- function(r, k, n){
  return((r*(1-min(n,k)/k))*min(n,k))
}
growth(r = 10, k = 100, n = 10)
```


Here we build the our simulator function using a while loop that will run the simulation until one
population 'wins'. We call all functions made above and loop over them for each patch and each population within a patch.
Before using _if_ and _ifelse_ statements to determine the winner and return the results of each simulation.

```{r THE SIMULATOR}
simulate <- function(world, params){

i <- 0
while(i<1000){
  
#growth on all patches
  for(patch in seq(1, length(world))){
    for(pop in seq(1, length(world[[patch]]))){
      curr_pop_name <- names(world[[patch]][pop])
      curr_pop <- world[[patch]][[curr_pop_name]]
      world[[patch]][[curr_pop_name]] <- world[[patch]][[curr_pop_name]] + 
        round(growth(r = params[[curr_pop_name]][["r"]], k = params[[curr_pop_name]][["k"]], n = curr_pop)) -> curr_pop
    }
  }
  
  #movement on all patches
  for(patch in seq(1, length(world))){
    for(pop in seq(1, length(world[[patch]]))){
      curr_pop_name <- names(world[[patch]][pop])
      curr_pop <- world[[patch]][[curr_pop_name]]
      migrants <- migrants_number(floor(curr_pop))
      direction <- direction_generator()
      destination <- patch + direction
      if(all(destination > 0 & destination <= length(world))){
        world[[patch]][[curr_pop_name]] <- world[[patch]][[curr_pop_name]] - migrants
        world[[destination]][[curr_pop_name]] <- world[[destination]][[curr_pop_name]] + migrants
    }
   }
  }
  
#checking if one of them won
  if(sum(unlist(world$"20")) > 0){
    result <- list(world)
    edge_patch <- world$"20"
  #checking which one said winner is
    the_winner <- character()
    ifelse(edge_patch$n1 > 0 && edge_patch$n2 == 0, the_winner <- "n1", 
           ifelse(edge_patch$n1 == 0 && edge_patch$n2 > 0, the_winner <- "n2", the_winner <- NA))
    
    #put the results in a new row in simulation_data
    
    simulation_data %>% add_row(worlds = result, timestep = i, parameters = list(params),
                                winner = the_winner,  r1 = params$n1$r, 
                                r2 = params$n2$r, k1 = params$n1$k, k2 = params$n2$k) -> simulation_data
    return(simulation_data)
  }
  i <- i + 1
}
  }
```

Making an empty data frame to store the runs of multiple simulation runs.
```{r clearing dataframe}
#making the dataframe empty
simulation_data <- tibble(worlds = list(), timestep = numeric(), parameters = list(),
    winner = character(), r1 = numeric(), r2 = numeric(), k1 = numeric(), k2 = numeric())

```

Here we loop over the simulation a 100 times at the same parameters.
```{r running the simulation}
for(i in seq(1, 100, 1)){
simulation_data <- simulate(my_world, my_params)
}
```

Plotting the results of our simulation in a histograms in order to look at the numbers of wins 
for each population at a given set of parameters
```{r 1 sim isolated}
  simulation_data %>% filter(!is.na(winner)) %>% 
  group_by(winner, r1, k1, r2, k2)  %>% 
  filter(r1 == 0.3, k1 == 300, r2 == 0.6, k2 == 150) %>% 
  ggplot(aes(x = timestep))+
  scale_fill_ordinal()+
  facet_wrap(~r1+r2+k1+k2, labeller = labeller(.multi_line = FALSE, .default = label_both))+
  geom_histogram(aes(x = timestep, fill = winner), binwidth = 1, colour = "black")+
  labs(x = "Win Time", y = "# of Wins", title = "Comparison of Win Speed at Different r and K Values")+
  theme(text = element_text(family = "Tahoma"), 
        axis.title = element_text(face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid = element_line(colour = "lightgrey"))
 
```

```{r combined}
  simulation_data %>% filter(!is.na(winner)) %>% group_by(winner, r1, k1, r2, k2) %>% 
  ggplot(aes(x = timestep))+
  scale_fill_ordinal()+
  facet_wrap(~r1+r2+k1+k2, labeller = labeller(.multi_line = FALSE, .default = label_both))+
  geom_histogram(aes(x = timestep, fill = winner), binwidth = 1, colour = "black")+
  labs(x = "Win Time", y = "# of Wins", title = "Comparison of Win Speed at Different r and K Values")+
  theme(text = element_text(family = "Tahoma"), 
        axis.title = element_text(face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid = element_line(colour = "lightgrey"))
 
```
