

Just loading any necessary libraries

```{r libraries}
library(tidyverse)
library(ggplot2)
```

For proof-of-concept we are currently manually creating one specific world and set of parameters to demonstrate how the functions will run on this example
```{r worlds and params generator}
my_world <- list("1" = list("n1" = 10, "n2" = 10), 
              "2" = list("n1" = 0, "n2" = 0), 
              "3" = list("n1" = 0, "n2" = 0),
              "4" = list("n1" = 0, "n2" = 0), 
              "5" = list("n1" = 0, "n2" = 0), 
              "6" = list("n1" = 0, "n2" = 0), 
              "7" = list("n1" = 0, "n2" = 0), 
              "8" = list("n1" = 0, "n2" = 0), 
              "9" = list("n1" = 0, "n2" = 0), 
              "10" = list("n1" = 0, "n2" = 0),
              "11" = list("n1" = 0, "n2" = 0),
              "12" = list("n1" = 0, "n2" = 0),
              "13" = list("n1" = 0, "n2" = 0),
              "14" = list("n1" = 0, "n2" = 0),
              "15" = list("n1" = 0, "n2" = 0),
              "16" = list("n1" = 0, "n2" = 0),
              "17" = list("n1" = 0, "n2" = 0),
              "18" = list("n1" = 0, "n2" = 0),
              "19" = list("n1" = 0, "n2" = 0),
              "20" = list("n1" = 0, "n2" = 0))
my_params <- list("n1" = list("r" = 0.9, "k" = 900), 
               "n2" = list("r" = 0.1, "k" = 100))
```
Now that the world is created, we have made a set of functions and algorithms to run our simulation.


The following will determine how many individuals will migrate from a patch based on a binomial draw
```{r migrants generator}
migrants_number <- function(n){
  return(rbinom(n = 1, size = n, prob = 0.5))
}
```

The following will randomly determine which direction a population will migrate to.
```{r direction generator function}
direction_generator <-function(){
  return(sample(c(-1, 1), 1))
}
direction_generator()

```

The following is the population growth, that calculates by how much a given population will grow based on the logistic growth equation 
```{r growth function}
growth <- function(r, k, n){
  return((r*(1-min(n,k)/k))*min(n,k))
}
growth(r = 10, k = 100, n = 10)
```


The following chunk simulates the the growth and migration for 10 timesteps, note that there are a lot of things that we have hard coded a lot of the steps here instead of creating functions for everything so that we have a proof of concept that works. It first applies the logistic growth rate on all the populations in all the patches, then applies the movement function on each population in each patch after they have grown for 1 timestep, as this meets our assumptions more accurately.
the plan is to have each of these steps converted to their unique functions of growth and movement across any world of any size. Then they will be combined to one "simulate" function so that eventually we can include the checker that will stop the simulation when one of the populations wins. 

please ignore the comments which were used for troubleshooting

also note that currently the _ceiling()_ function is being called on the n parameter when determining the number of migrants, as not doing so causes rbinom to take a float, that was generated as a result of us using the logistic growth function, and rbinom cannot work on floats, it requires whole numbers/integers
```{r THE SIMULATOR}
simulate <- function(world, params){

i <- 0
while(i<1000){
  
#growth on all patches
  for(patch in seq(1, length(world))){
    for(pop in seq(1, length(world[[patch]]))){
      curr_pop_name <- names(world[[patch]][pop])
      curr_pop <- world[[patch]][[curr_pop_name]]
      world[[patch]][[curr_pop_name]] <- world[[patch]][[curr_pop_name]] + round(growth(r = params[[curr_pop_name]][["r"]], k = params[[curr_pop_name]][["k"]], n = curr_pop)) -> curr_pop
    }
  }
  
  #movement on all patches
  for(patch in seq(1, length(world))){
    for(pop in seq(1, length(world[[patch]]))){
      curr_pop_name <- names(world[[patch]][pop])
      curr_pop <- world[[patch]][[curr_pop_name]]
      migrants <- migrants_number(floor(curr_pop))
      direction <- direction_generator()
      destination <- patch + direction
      if(all(destination > 0 & destination <= length(world))){
        world[[patch]][[curr_pop_name]] <- world[[patch]][[curr_pop_name]] - migrants
        world[[destination]][[curr_pop_name]] <- world[[destination]][[curr_pop_name]] + migrants
    }
   }
  }
  
#checking if one of them won
  if(sum(unlist(world$"20")) > 0){
    result <- list(world)
    edge_patch <- world$"20"
  #checking which one said winner is
    the_winner <- character()
    ifelse(edge_patch$n1 > 0 && edge_patch$n2 == 0, the_winner <- "n1", ifelse(edge_patch$n1 == 0 && edge_patch$n2 > 0, the_winner <- "n2", the_winner <- NA))
    
    #put the results in a new row in simulation_data
    
    simulation_data %>% add_row(worlds = result, timestep = i, parameters = list(params), winner = the_winner,  r1 = params$n1$r, r2 = params$n2$r, k1 = params$n1$k, k2 = params$n2$k) -> simulation_data
    return(simulation_data)
  }
  i <- i + 1
}
  }
```

```{r clearing dataframe}
#making the dataframe empty
simulation_data <- tibble(worlds = list(), timestep = numeric(), parameters = list(), winner = character(), r1 = numeric(), r2 = numeric(), k1 = numeric(), k2 = numeric())

```


```{r running the simulation}
for(i in seq(1, 900, 1)){
simulation_data <- simulate(my_world, my_params)
}
```

```{r 1 sim isolated}
  simulation_data %>% filter(!is.na(winner)) %>% group_by(winner, r1, k1, r2, k2) %>% filter(r1 == 0.5, k1 == 900, r2 == 0.5, k2 == 100) %>% 
  ggplot(aes(x = timestep))+
  scale_fill_ordinal()+
  facet_wrap(~r1+r2+k1+k2, labeller = labeller(.multi_line = FALSE, .default = label_both))+
  geom_histogram(aes(x = timestep, fill = winner), binwidth = 1, colour = "black")+
  labs(x = "Win Time", y = "# of Wins", title = "Comparison of Win Speed at Different r and K Values")+
  theme(text = element_text(family = "Tahoma"), 
        axis.title = element_text(face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid = element_line(colour = "lightgrey"))
 
```

```{r combined}
  simulation_data %>% filter(!is.na(winner)) %>% group_by(winner, r1, k1, r2, k2) %>% 
  ggplot(aes(x = timestep))+
  scale_fill_ordinal()+
  facet_wrap(~r1+r2+k1+k2, labeller = labeller(.multi_line = FALSE, .default = label_both))+
  geom_histogram(aes(x = timestep, fill = winner), binwidth = 1, colour = "black")+
  labs(x = "Win Time", y = "# of Wins", title = "Comparison of Win Speed at Different r and K Values")+
  theme(text = element_text(family = "Tahoma"), 
        axis.title = element_text(face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid = element_line(colour = "lightgrey"))
 
```
