

Just loading any necessary libraries

```{r libraries}
# Loading all necessary libraries
library(tidyverse)
library(ggplot2)
library(extrafont)#just extra fonts for plotting
```

To create our world, we made a nested list containing all the patches and the populations for both n1 and n2 inside. With a separate nested list containing the parameters for both populations.
```{r worlds and params generator}
my_world <- list("1" = list("n1" = 10, "n2" = 10), #creating the world, which includes 20 patches of 2 populations
              "2" = list("n1" = 0, "n2" = 0), 
              "3" = list("n1" = 0, "n2" = 0),
              "4" = list("n1" = 0, "n2" = 0), 
              "5" = list("n1" = 0, "n2" = 0), 
              "6" = list("n1" = 0, "n2" = 0), 
              "7" = list("n1" = 0, "n2" = 0), 
              "8" = list("n1" = 0, "n2" = 0), 
              "9" = list("n1" = 0, "n2" = 0), 
              "10" = list("n1" = 0, "n2" = 0),
              "11" = list("n1" = 0, "n2" = 0),
              "12" = list("n1" = 0, "n2" = 0),
              "13" = list("n1" = 0, "n2" = 0),
              "14" = list("n1" = 0, "n2" = 0),
              "15" = list("n1" = 0, "n2" = 0),
              "16" = list("n1" = 0, "n2" = 0),
              "17" = list("n1" = 0, "n2" = 0),
              "18" = list("n1" = 0, "n2" = 0),
              "19" = list("n1" = 0, "n2" = 0),
              "20" = list("n1" = 0, "n2" = 0))
my_params <- list("n1" = list("r" = 0.6, "k" = 300), "n2" = list("r" = 0.6, "k" = 300)) #creating the parameters(set to one specific value we can change as we desire)
```
Now that the world is created, we  made a set of functions and algorithms to run our simulation

The migration function uses a random binomial draw to determine the number of migrants in a 
population
```{r migrants generator}
migrants_number <- function(number){# defining the function
  return(rbinom(n = 1, size = number, prob = 0.5))# returning a single random binomial sample, set at probability 0.5 from the population.
  #i.e this will return a binomially distributed random number from 0 -> the given populations number
}

```

The direction generator function random picks between 1 and -1 to determine what direction the
migrants will move.
```{r direction generator function}
direction_generator <-function(){# function definition
  return(sample(c(-1, 1), 1))# randomly returning + or - 1
}
direction_generator()#simply testing the function

```

Our popultion growth function uses a discrete time logsistic growth equation, that produces the 
population in the next time step instead of returning a rate of change.
```{r growth function}
growth <- function(r, k, n){# function definition
  return((r*(1-min(n,k)/k))*min(n,k))# returning the n(t+1) of discrete time logistic growth.
}
growth(r = 10, k = 1000, n = 10)
```


Here we build the our simulator function using a while loop that will run the simulation until one
population 'wins'. We call all functions made above and loop over them for each patch and each population within a patch.
Before using _if_ and _ifelse_ statements to determine the winner and return the results of each simulation.

```{r THE SIMULATOR}
simulate <- function(world, params){ #function definition

i <- 0 #i counter creaion
while(i<1000){ # while loop which stops when i reaches 1000 (there's a i += 1 below)
#growth on all patches
  for(patch in seq(1, length(world))){
    for(pop in seq(1, length(world[[patch]]))){# iterating through each population in each patch
      curr_pop_name <- names(world[[patch]][pop])# isolating the iterated population name for easier use
      curr_pop <- world[[patch]][[curr_pop_name]]# isolating the iterated population value for easier use
      world[[patch]][[curr_pop_name]] <- world[[patch]][[curr_pop_name]] + 
        round(growth(r = params[[curr_pop_name]][["r"]], k = params[[curr_pop_name]][["k"]], n = curr_pop)) -> curr_pop# calculating the population at the next timestep, and adding that it to curr_pop
    }
  }
# so far the logistic growth step of our system's life cycle was applied to every population in every patch

#next we will apply the migration
  
  #movement on all patches
  for(patch in seq(1, length(world))){
    for(pop in seq(1, length(world[[patch]]))){# iterating through each population in each patch
      curr_pop_name <- names(world[[patch]][pop])# isolating the iterated population name for easier use
      curr_pop <- world[[patch]][[curr_pop_name]]# isolating the iterated population value for easier use
      migrants <- migrants_number(floor(curr_pop))# determining the number of migrants
      direction <- direction_generator()# determining which direction the will go
      destination <- patch + direction# isolating the destination patch
      if(all(destination > 0 & destination <= length(world))){# checking if the destination patch exists
        world[[patch]][[curr_pop_name]] <- world[[patch]][[curr_pop_name]] - migrants# subtracting the migrants from original patch
        world[[destination]][[curr_pop_name]] <- world[[destination]][[curr_pop_name]] + migrants# adding the migrants into new patch
    }
   }
  }
  
#checking if one of the populations won
  if(sum(unlist(world$"20")) > 0){# checking if the final patch has any members in it( = 0 if empty, > 0 otherwise). If it does:
    result <- list(world)# save the state of the world, as is into result
    edge_patch <- world$"20"# Isolate the final patch so that we can:
  #check which one the winner is
    the_winner <- character()
    ifelse(edge_patch$n1 > 0 && edge_patch$n2 == 0, the_winner <- "n1", 
           ifelse(edge_patch$n1 == 0 && edge_patch$n2 > 0, the_winner <- "n2", the_winner <- NA))# isolate the winner based on the ifelse statement
    
    #put the results in a new row in simulation_data
    
    simulation_data %>% add_row(worlds = result, timestep = i, parameters = list(params),
                                winner = the_winner,  r1 = params$n1$r, 
                                r2 = params$n2$r, k1 = params$n1$k, k2 = params$n2$k) -> simulation_data
    return(simulation_data)#retun the data as a whole
  }
  i <- i + 1
}
  }
```

Making an empty data frame to store the runs of multiple simulation runs.
```{r clearing dataframe}
#making the data frame (empty at first)
simulation_data <- tibble(worlds = list(), timestep = numeric(), parameters = list(),
    winner = character(), r1 = numeric(), r2 = numeric(), k1 = numeric(), k2 = numeric())

```

Here we loop over the simulation a 100 times at the same parameters.
```{r running the simulation}
for(i in seq(1, 400, 1)){# for loop of 400 iterations, to give us 400 replicates of the simulation
simulation_data <- simulate(my_world, my_params)# running the simulation (once)
}
```


```{r combined}
  simulation_data %>% 
  filter(!is.na(winner)) %>% #removing any NA points (the tie cases)
  group_by(winner, r1, k1, r2, k2) %>%# arranging the data neatly across the columns
  ggplot(aes(x = timestep, colour = winner))+
  facet_wrap(~r1+r2+k1+k2, labeller = labeller(.multi_line = FALSE, .default = label_both))+# faceting over the parameters then labeling them appropriately
  geom_freqpoly(data = subset(simulation_data, winner == "n2"), binwidth = 1)+#the frequency polygon for whenever n2 won
  geom_freqpoly(data = subset(simulation_data, winner == "n1"), binwidth = 1,  aes(y = ..count..*(1)))+# frequency polygon for whenever n1 has won(the y aesthetic is there to allow us to invert the line for comparison)
  labs(x = "Win Time", y = "# of Wins", title = "Comparison of Win Speed at Different r and K Values")+# labelling
  theme(text = element_text(family = "Tahoma"),# stylistic changes
        axis.title = element_text(face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid = element_line(colour = "transparent"))
 
```
